\documentclass{beamer}

\usepackage{listings}

\usetheme{default}

\title{Proof of Regular Expression Matching CPS Algorithm Termination in Agda}

\author{Joomy Korkut \and Maksim Trifunovski \and Dan Licata}

\institute {
Department of Mathematics and Computer Science\\
Wesleyan University
}

\date{June 24th, 2015}

\begin{document}

\begin{frame}
\titlepage
\end{frame}

\begin{frame}{Outline}
\tableofcontents
\end{frame}

\section{Definitions}

\subsection{Agda}

\begin{frame}{Agda}
\begin{itemize}
\item Agda is a dependently typed functional programming language mostly used to write proofs.
\item Agda is a \emph{total} language, which means all functions in Agda must terminate and return a value.
\end{itemize}
\end{frame}

\subsection{Regular Expressions}

\begin{frame}{Regular Expressions}
\begin{itemize}
\item Regular expressions (regex) are a way of checking for patterns and extracting information from sequences of characters.
\item Examples include: checking if an e-mail address is valid, finding the type of a credit card number.
\end{itemize}
\end{frame}

\begin{frame}{Definition of Regular Expressions}
\begin{itemize}
\item $\emptyset$ : The empty set.
\item $\varepsilon$ : The empty string.
\item $a$ : The character literal, if $a$ is in our alphabet ($a \in \Sigma$).
\item Suppose $r_1, r_2$ are regular expressions.\\
$r_1 + r_2$ : Alteration. (matches $r_1$ or $r_2$)
\item Suppose $r_1, r_2$ are regular expressions.\\
$r_1 r_2$ : Concatenation. (a prefix matches $r_1$ and the following suffix matches $r_2$)
\item Suppose $r$ is a regular expression.\\
$r^*$ : Kleene star. (matches $r$ zero or more times)
\end{itemize}
\begin{example}
$(a + ... + z + 0 + 1 + ... + 9)^* @ (a + ... + z + 0 + 1 + ... + 9)^* . (com + net + org + edu + ...)$
\end{example}
\end{frame}

\subsection{Continuation Passing Style}

\begin{frame}{Continuation Passing Style}
\begin{itemize}
\item The first matching function that comes to mind would be \\
$\text{match} : \text{RegExp} \to \text{List Char} \to \text{Bool}$

However, in the $r_1 r_2$ case we have to check that there exists a split of the string $s = s_1 s_2$, such that $s_1 \in L(r_1)$ and $s_2 \in L(r_2)$.

\item Hence, we make a generalization such that our function matches a prefix of the char list, calls the continuation function with the rest of the list and return the result. \\
$\text{match} : \text{RegExp} \to \text{List Char} \to (\text{List Char} \to \text{Bool}) \to \text{Bool}$

\item Overall, if match r s $(\lambda x \to x == [])$ returns true, then the regex accepts our string.
\end{itemize}
\end{frame}

\section{Our Research}

\subsection{Termination Proof}

\begin{frame}{Termination Problem}
\begin{itemize}
\item  In the case of the regex $r^*$, since the regex accepts any number of strings that match $r$, including $0$, the string that we pass to the continuation function might stay the same as the original string.
\item Agda doesn't believe that this function will always terminate. In order to create a workaround, we used regexes in standard form, such that $L(standardize(r)) = L(r) - \{\varepsilon\}$.
\item In standard form, $r^+$ can only accept one or more number of strings that match $r$. Hence the continuation function always passes a smaller string than the string we started with and therefore Agda believes that the program terminates.
\end{itemize}
\end{frame}

\subsection{Soundness, Completeness and Future Work}

\begin{frame}{Soundness, Completeness and Future Work}
We are now working on the soundness and completeness of the algorithm, and we are mostly done.
\begin{itemize}
\item Soundness: Show that if match $r$ $s$ $k$ is true, then there is a split of $s$, namely $s_1 s_2$, such that $s_1 \in L(r)$ and $k$ $s_2$ is true.
\item Completeness: Show that if there is a split of $s$, namely $s_1 s_2$, such that $s_1 \in L(r)$ and $k$ $s_2$ is true, then match $r$ $s$ $k$ is true.
\item Standardization: Our next step is to show that for any regex $r$, $L(r) = L(standardize(r)) \cup \delta(r)$, where $\delta(r) = \varepsilon$ if $r$ accepts the empty string, $\emptyset$ otherwise.
\end{itemize}
\end{frame}

\begin{frame}{Code}
The Agda code is available at\\
\url{http://github.com/joom/regexp-agda}
\end{frame}

\end{document}
